---
title: "Regressions"
author: "Alejandra Garcia Isaza"
date: '2022-07-12'
output: html_document
---

Diss Research questions:

1. Can both structural and relational components of family-school connections be identified using exploratory factor analysis approaches in the Conexiones parents’ assessment data? --> addressed with Exploratory Factor Analyses, not included here.
    
    I identified *15 predictors* in the EFA part of my study. They are different parent engagement practices, some are considered structural     and some are considered relational ("hwi_1"    "hwi_2"    "hwi_3"      "monit"    "str_home" "convos_1" "convos_2" "convos_3" "sbi_1"        "sbi_2"    "belong"  "endorse"  "value_ed" "rel_1"    "rel_2" )
    
    The *outcome* is youth engagement (called "engage")
    
    The *moderators* are parent's educational level (q36_hv1) and parent's comfort with ENG (q173_3_p1)

2. What specific structural and relational components of family-school connections in the Conexiones parents’ assessment data are significantly and positively associated with positive students’ school engagement? --> to be addressed with simple linear regression models

3. Does parent educational level and comfort with English language moderate the relationship between structural and relational family-school connections components and student’s school engagement? --> to be addressed with moderated multiple regression models

4. What model best predicts positive students’ school engagement, a structural components only model or a combined structural plus relational components model? --> to be addressed with a step-wise multiple linear regression model, with predictor blocks


----
Questions for Jeff:
----

- Is it okay to treat my predictors and outcome (that really are Likert scales = ordinal variables) as numeric (dbl in R)? I didn't in EFAs because it was against EFA best practices. Not sure about rec. approach. 
- The moderators are parent's educational level and comfort with ENG. Should I treat these as factors?
- Should I include covariates in all the regression models (simple, multiple, etc)? 
- Was correct to conduct 15 separate simple regression models or should have I done it in one call? --> I want to know if the specific var predicts the outcome (RQ1)
- When conducting simple regression the p. value is the same in two places in the output, but when conducting in moderation, shows different values in the same two places. Why?
- For moderation need to grand mean center? (If yes, my code to do this on lines 110-137 is not working, any clues why?)
- Should I run moderation models if the IV and DV are not significantly related in simple linear regression models (models 11-15 below that have relational parent practice


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)
library(MVN)

theme_set(theme_minimal())
```


# Loading the dataset (only analysis vars)
```{r}
d <- read_sav(here("data", "d3.sav"))

view_df(d)
```


```{r}
d
```
# Scale creation (15 predictors, 1 outcome)
```{r}
d_w_scales <- d %>%
  rowwise() %>% 
  mutate(hwi_1 = mean(c(q86_p1, q87_p1, q88_p1, q89_p1, q94_p1, q96_p1, q97_p1), na.rm = TRUE),
         hwi_2 = mean(c(q84_p1, q85_p1, q93_p1, q95_p1), na.rm = TRUE),
         hwi_3 = mean(c(q90_p1, q91_p1, q98_p1), na.rm = TRUE),
         monit = mean(c(q137_p1, q139_p1, q140_p1, q143_p1, q144_p1), na.rm = TRUE),
         str_home = mean(c(q23_p1, q24_p1, q25_p1, q26_p1, q27_p1, q28_p1, q29_p1, q30_p1), na.rm = TRUE),
         convos_1 = mean(c(q76_p1, q77_p1, q78_p1, q79_p1), na.rm = TRUE),
         convos_2 = mean(c(q73_p1, q75_p1, q80_p1, q81_p1), na.rm = TRUE),
         convos_3 = mean(c(q68_p1, q69_p1, q70_p1, q71_p1), na.rm = TRUE),
         sbi_1 = mean(c(q31_p1, q32_p1, q33_p1, q34_p1), na.rm = TRUE),
         sbi_2 = mean(c(q36_p1, q37_p1, q38_p1, q39_p1, q40_p1), na.rm = TRUE),
         belong = mean(c(q47_p1, q48_p1, q51_p1, q52_p1, q53_p1), na.rm = TRUE),
         endorse = mean(c(q54_p1, q56_p1, q57_p1), na.rm = TRUE),
         value_ed = mean(c(q105_p1, q106_p1, q107_p1, q108_p1, q109_p1, q110_p1), na.rm = TRUE),
         rel_1 = mean(c(q41_p1, q58_p1, q62_p1, q63_p1, q64_p1, q65_p1, q66_p1, q67_p1), na.rm = TRUE),
         rel_2 = mean(c(q43_p1, q44_p1, q50_p1, q126_p1), na.rm = TRUE),
         engage = mean(c(q83_y1, q85_y1, q86_y1, q87_y1, q88_y1, q89_y1, q90_y1, q91_y1), na.rm = TRUE))

view_df(d_w_scales)
```

# descriptive stats function
```{r include=FALSE}

# function that takes a df and returns a df with only the numeric columns
only_numeric <- function(df) {
  select_numeric <- dplyr::select_if(df, is.numeric)
  return(select_numeric)
}

# a list with the functions we want
smry <- list(M = function(x) round(mean(x, na.rm = TRUE), 2), 
             SD = function(x) round(sd(x, na.rm = TRUE), 2),
             Min = function(x) round(min(x, na.rm = TRUE), 2),
             Max = function(x) round(max(x, na.rm = TRUE), 2))

# wrapping solution in a function
descriptives <- function(df) {
  select_numeric <- only_numeric(df)
  mean_sd <- map_df(select_numeric, function(col) map_df(smry, ~.x(col)),
       .id = "column")
  return(mean_sd)
}
```


```{r}
only_scales <- d_w_scales %>%
  select(108:123)

descrip_table <- descriptives(only_scales) %>%
  rename(c("Scale" = "column"))
```


# grand mean center predictors (?)
```{r}
d_w_scales_c <- d_w_scales %>%
mutate(hwi_1_c = hwi_1 - (round(mean(hwi_1, na.rm = TRUE), 2)),
       hwi_2_c = hwi_2 - mean(hwi_2, na.rm = TRUE),
       hwi_3_c = hwi_3 - mean(hwi_3, na.rm = TRUE),
       monit_c = monit - mean(monit, na.rm = TRUE),
       str_home_c = str_home - mean(str_home, na.rm = TRUE),
       convos_1_c = convos_1 - mean(convos_1, na.rm = TRUE),
       convos_2_c = convos_2 - mean(convos_2, na.rm = TRUE),
       convos_3_c = convos_3 - mean(convos_3, na.rm = TRUE),
       sbi_1_c = sbi_1 - mean(sbi_1, na.rm = TRUE),
       sbi_2_c = sbi_2 - mean(sbi_2, na.rm = TRUE),
       belong_c = belong - mean(belong, na.rm = TRUE),
       endorse_c = endorse - mean(endorse, na.rm = TRUE),
       value_ed_c = value_ed - mean(value_ed, na.rm = TRUE),
       rel_1_c = rel_1 - mean(rel_1, na.rm = TRUE),
       rel_2_c = rel_2 - mean(rel_2, na.rm = TRUE),
       ed_level = factor(q36_hv1),
       eng_comf = factor(q173_3_p1)) # this is not working. it gives me a zero in all the centered vars. Have tried multiple ways of using ()

# summary(d_w_scales)

# first row: 1.857143 - 2.704045 # -0.846902\
d_w_scales$hwi_1 - round(mean(d_w_scales$hwi_1, na.rm = TRUE), 2) # this works

d_w_scales_c <- d_w_scales %>%
  mutate(hwi_1_c = hwi_1 - round(mean(hwi_1, na.rm = TRUE), 2)) # this doesn't, first row of hwi_1_c = -0.002857143
```


#correlation matrix 
```{r}
library(Hmisc)
rcorr(as.matrix(only_scales), type = "pearson")
```

------------------------------
# RESEARCH QUESTION 2
------------------------------

# 2.1. Academic approach (tutor)
```{r}
m_hwi_1 <- lm(engage ~ hwi_1, d_w_scales)

summary(m_hwi_1) # p = 0.07
```
engage predicted by hw1

Intercept: expected value of youth's engagement when the average hW involvement in data set is considered (I think = 2.70)

slope: the effect HW involvement has in youth's engagement = for every one unit increase in HW involvement, youth's engagement increases by 0.15 (but is not significant).

# 2.2. structuring/monitoring approach to HWI
```{r}
m_hwi_2 <- lm(engage ~ hwi_2, d_w_scales)

summary(m_hwi_2)# p = 0.01
```

# 2.3. positive reinforcement to increase HW completion
```{r}
m_hwi_3 <- lm(engage ~ hwi_3, d_w_scales)

summary(m_hwi_3) # p = 0.03
```

# 2.4. monitoring
```{r}
m_monit <- lm(engage ~ monit, d_w_scales)

summary(m_monit) # p = 0.89
```


# 2.5. structure at home
```{r}
m_str_home <- lm(engage ~ str_home, d_w_scales)

summary(m_str_home) # p = 0.04
```


# 2.6. convos: child beh & school climate
```{r}
m_convos_1 <- lm(engage ~ convos_1, d_w_scales)

summary(m_convos_1) # p = 0.86
```


# 2.7. convos: future career planning
```{r}
m_convos_2 <- lm(engage ~ convos_2, d_w_scales)

summary(m_convos_2) # p = 0.07
```


# 2.8. convos: child's social involvement
```{r}
m_convos_3 <- lm(engage ~ convos_3, d_w_scales)

summary(m_convos_3) # p = 0.27
```


# 2.9. Basic involvement
```{r}
m_sbi_1 <- lm(engage ~ sbi_1, d_w_scales)

summary(m_sbi_1) # p = 0.05
```


# 2.10. Resource-intensive involvement
```{r}
m_sbi_2 <- lm(engage ~ sbi_2, d_w_scales)

summary(m_sbi_2) # p = 0.03
```


# 2.11. Belongingness
```{r}
m_belong <- lm(engage ~ belong, d_w_scales)

summary(m_belong) # p = 0.97
```


# 2.12. Endorsemnt
```{r}
m_endorse <- lm(engage ~ endorse, d_w_scales)

summary(m_endorse) # p = 0.53
```


# 2.13. Value & supp of Ed
```{r}
m_value_ed <- lm(engage ~ value_ed, d_w_scales)

summary(m_value_ed) # p = 0.77
```


# 2.14. Rel with teachers: comfort, interest, and openness
```{r}
m_rel_1 <- lm(engage ~ rel_1, d_w_scales)

summary(m_rel_1) # p = 0.19
```


# 2.15. Rel with school: parent self-efficacy to solve probs with school
```{r}
m_rel_2 <- lm(engage ~ rel_2, d_w_scales)

summary(m_rel_2) # p = 0.65
```

# just for fun: relational aspects predicting parent's self-effi
```{r}
m_rel_belong <- lm(rel_2 ~ belong, d_w_scales)
summary(m_rel_belong) # p < 0.001

m_rel_endorse <- lm(rel_2 ~ endorse, d_w_scales)
summary(m_rel_endorse) # p < 0.001

m_rel_value_ed <- lm(rel_2 ~ value_ed, d_w_scales)
summary(m_rel_value_ed) # p < 0.001

m_rel_qual <- lm(rel_2 ~ rel_1, d_w_scales)
summary(m_rel_qual) # p < 0.001
```


------------------------------
# RESEARCH QUESTION 3
------------------------------

```{r}
d_w_scales$q36_hv1 # ed level ((1 - 11)

d_w_scales$q173_3_p1 # ENG comfort in youth school (1 - 5)
```


# 3.1. Academic approach (tutor)
```{r}
m2.1_hwi_1 <- lm(engage ~ hwi_1 + hwi_1:q36_hv1, d_w_scales)
summary(m2.1_hwi_1) # p = not sure which to look at

m2.2_hwi_1 <- lm(engage ~ hwi_1 + hwi_1:q173_3_p1, d_w_scales)
summary(m2.2_hwi_1) # p = not sure which to look at
```


# 3.2. structuring/monitoring approach to HWI
```{r}
m2.1_hwi_2 <- lm(engage ~ hwi_2 + hwi_2:q36_hv1, d_w_scales)
summary(m2.1_hwi_2) # p = 

m2.2_hwi_2 <- lm(engage ~ hwi_2 + hwi_2:q173_3_p1, d_w_scales)
summary(m2.2_hwi_2) # p = 
```

## Note for Jeff: 
disregard the following models, I just copied and pasted the simple regression models and will add the interaction terms once I get my mind sorted out. 

# 3.3. positive reinforcement to increase HW completion
```{r}
m_hwi_3 <- lm(engage ~ hwi_3, d_w_scales)

summary(m_hwi_3) # p = 0.03
```

# 3.4. monitoring
```{r}
m_monit <- lm(engage ~ monit, d_w_scales)

summary(m_monit) # p = 0.89
```


# 3.5. structure at home
```{r}
m_str_home <- lm(engage ~ str_home, d_w_scales)

summary(m_str_home) # p = 0.04
```


# 3.6. convos: child beh & school climate
```{r}
m_convos_1 <- lm(engage ~ convos_1, d_w_scales)

summary(m_convos_1) # p = 0.86
```


# 3.7. convos: future career planning
```{r}
m_convos_2 <- lm(engage ~ convos_2, d_w_scales)

summary(m_convos_2) # p = 0.07
```


# 3.8. convos: child's social involvement
```{r}
m_convos_3 <- lm(engage ~ convos_3, d_w_scales)

summary(m_convos_3) # p = 0.27
```


# 3.9. Basic involvement
```{r}
m_sbi_1 <- lm(engage ~ sbi_1, d_w_scales)

summary(m_sbi_1) # p = 0.05
```


# 3.10. Resource-intensive involvement
```{r}
m_sbi_2 <- lm(engage ~ sbi_2, d_w_scales)

summary(m_sbi_2) # p = 0.03
```


# 3.11. Belongingness
```{r}
m_belong <- lm(engage ~ belong, d_w_scales)

summary(m_belong) # p = 0.97
```


# 3.12. Endorsemnt
```{r}
m_endorse <- lm(engage ~ endorse, d_w_scales)

summary(m_endorse) # p = 0.53
```


# 3.13. Value & supp of Ed
```{r}
m_value_ed <- lm(engage ~ value_ed, d_w_scales)

summary(m_value_ed) # p = 0.77
```


# 3.14. Rel with teachers: comfort, interest, and openness
```{r}
m_rel_1 <- lm(engage ~ rel_1, d_w_scales)

summary(m_rel_1) # p = 0.19
```


# 3.15. Rel with school: parent self-efficacy to solve probs with school
```{r}
m_rel_2 <- lm(engage ~ rel_2, d_w_scales)

summary(m_rel_2) # p = 0.65
```





