---
title: "regression model diagnostics"
author: "Alejandra Garcia Isaza"
date: '2022-07-19'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rio)
library(here)
library(dplyr)
library(haven)

theme_set(theme_minimal())
```

```{r}
d <- read_sav(here("data", "d4 - scales & covariates.sav"))
```



**Model assumptions & diagnostics** 


# Normality assumption: outcome
```{r, engage}
d$engage

ggplot(d, aes(engage)) +
geom_histogram(alpha = 0.7) # meets normality assumption? assumption may be tenable, maybe presence of an outlier (score 2.250) participant row 81

d_no81 <- d %>%
  filter(participant_id != "608_1") # 93 rows now

ggplot(d_no81, aes(engage)) +
  geom_histogram(alpha = 0.7)
                    
```

1. Academic approach (tutor)

# linearity
```{r}
ggplot(d, aes(hwi_1, engage)) +
geom_point() +
geom_smooth() +
geom_smooth(method = "lm",
color = "magenta")

ggplot(d_no81, aes(hwi_1, engage)) +
geom_point() +
geom_smooth() +
geom_smooth(method = "lm",
color = "magenta")
```
# plots - univariate
```{r}
m_hwi_1a <- lm(engage ~ hwi_1, d)
plot(m_hwi_1a)

m_hwi_1a_d_no81 <- lm(engage ~ hwi_1, d_no81)
plot(m_hwi_1a_d_no81)
```
```{r}
coef(m_hwi_1a)
coef(m_hwi_1a_d_no81)
```


# outliers
```{r}
plot(m_hwi_1a, which = 4)

plot(m_hwi_1a_d_no81, which = 4)
```


2. structuring/monitoring approach to HWI

# linearity
```{r}
ggplot(d, aes(hwi_2, engage)) +
geom_point() +
geom_smooth() +
geom_smooth(method = "lm",
color = "magenta")

ggplot(d_no81, aes(hwi_2, engage)) +
geom_point() +
geom_smooth() +
geom_smooth(method = "lm",
color = "magenta")
```
# plots
```{r}
m_hwi_2a <- lm(engage ~ hwi_2, d)
plot(m_hwi_2a)

m_hwi_2a_d_no81 <- lm(engage ~ hwi_2, d_no81)
plot(m_hwi_2a_d_no81)
```
```{r}
coef(m_hwi_2a)

coef(m_hwi_2a_d_no81)
```


# outliers
```{r}
plot(m_hwi_2a, which = 4)

plot(m_hwi_2a_d_no81, which = 4)
```


3. positive reinforcement to increase HW completion

# linearity
```{r}
ggplot(d, aes(hwi_3, engage)) +
geom_point() +
geom_smooth() +
geom_smooth(method = "lm",
color = "magenta")
```
# plots
```{r}
m_hwi_3a <- lm(engage ~ hwi_3, d)
plot(m_hwi_3a)

m_hwi_3a_d_no81 <- lm(engage ~ hwi_3, d_no81)
plot(m_hwi_3a_d_no81)
```
```{r}
coef(m_hwi_3a)

coef(m_hwi_3a_d_no81)
```

# outliers
```{r}
plot(m_hwi_3a, which = 4)

plot(m_hwi_3a_d_no81, which = 4)
```

4. monitoring

# linearity
```{r}
ggplot(d, aes(monit, engage)) +
geom_point() +
geom_smooth() +
geom_smooth(method = "lm",
color = "magenta")
```

# plots
```{r}
m_monit_a <- lm(engage ~ monit, d)
plot(m_monit_a)

m_monit_a_no81 <- lm(engage ~ monit, d_no81)
plot(m_monit_a_no81)
```

```{r}
coef(m_monit_a)

coef(m_monit_a_no81)
```

# outliers
```{r}
plot(m_monit_a, which = 4)

plot(m_monit_a_no81, which = 4)
```
5. structure at home

# linearity
```{r}
ggplot(d, aes(str_home, engage)) +
geom_point() +
geom_smooth() +
geom_smooth(method = "lm",
color = "magenta")
```
# plots
```{r}
m_str_home_a <- lm(engage ~ str_home, d)
plot(m_str_home_a)

m_str_home_a_no81 <- lm(engage ~ str_home, d_no81)
plot(m_str_home_a_no81)
```

```{r}
coef(m_str_home_a)

coef(m_str_home_a_no81)
```

# outliers
```{r}
plot(m_str_home_a, which = 4)

plot(m_str_home_a_no81, which = 4)
```
6. convos: child beh & school climate

# linearity
```{r}
ggplot(d, aes(convos_1, engage)) +
geom_point() +
geom_smooth() +
geom_smooth(method = "lm",
color = "magenta")
```

# plots
```{r}
m_convos_1a <- lm(engage ~ convos_1, d)
plot(m_convos_1a)

m_convos_1a_no81 <- lm(engage ~ convos_1, d_no81)
plot(m_convos_1a_no81)
```
```{r}
coef(m_convos_1a)

coef(m_convos_1a_no81)
```


# outliers
```{r}
plot(m_convos_1a, which = 4)

plot(m_convos_1a_no81, which = 4)
```
7. convos: future career planning

# linearity
```{r}
ggplot(d, aes(convos_2, engage)) +
geom_point() +
geom_smooth() +
geom_smooth(method = "lm",
color = "magenta")
```
# plots
```{r}
m_convos_2a <- lm(engage ~ convos_2, d)
plot(m_convos_2a)

m_convos_2a_no81 <- lm(engage ~ convos_2, d_no81)
plot(m_convos_2a_no81)
```
```{r}
coef(m_convos_2a)

coef(m_convos_2a_no81)
```

# outliers
```{r}
plot(m_convos_2a, which = 4)

plot(m_convos_2a_no81, which = 4)
```

8. convos: child's social involvement
# linearity
```{r}
ggplot(d, aes(convos_3, engage)) +
geom_point() +
geom_smooth() +
geom_smooth(method = "lm",
color = "magenta")
```

# plots
```{r}
m_convos_3a <- lm(engage ~ convos_3, d)
plot(m_convos_3a)

m_convos_3a_no81 <- lm(engage ~ convos_3, d_no81)
plot(m_convos_3a_no81)
```

```{r}
coef(m_convos_3a)

coef(m_convos_3a_no81)
```

# outliers
```{r}
plot(m_convos_3a, which = 4)

plot(m_convos_3a_no81, which = 4)
```
9. Basic involvement
# linearity
```{r}
ggplot(d, aes(sbi_1, engage)) +
geom_point() +
geom_smooth() +
geom_smooth(method = "lm",
color = "magenta")
```

# plots
```{r}
m_sbi_1a <- lm(engage ~ sbi_1, d)
plot(m_sbi_1a) 

m_sbi_1a_no81 <- lm(engage ~ sbi_1, d_no81)
plot(m_sbi_1a_no81) 
```
```{r}
coef(m_sbi_1a)

coef(m_sbi_1a_no81)
```


# outliers
```{r}
plot(m_sbi_1a, which = 4)

plot(m_sbi_1a_no81, which = 4)
```
10. Resource-intensive involvement
# linearity
```{r}
ggplot(d, aes(sbi_2, engage)) +
geom_point() +
geom_smooth() +
geom_smooth(method = "lm",
color = "magenta")
```

# plots
```{r}
m_sbi_2a <- lm(engage ~ sbi_2, d)
plot(m_sbi_2a) 

m_sbi_2a_no81 <- lm(engage ~ sbi_2, d_no81)
plot(m_sbi_2a_no81) 
```
```{r}
coef(m_sbi_2a)

coef(m_sbi_2a_no81)
```


# outliers
```{r}
plot(m_sbi_2a, which = 4)

plot(m_sbi_2a_no81, which = 4)
```
11. Belongingness
# linearity
```{r}
ggplot(d, aes(belong, engage)) +
geom_point() +
geom_smooth() +
geom_smooth(method = "lm",
color = "magenta")
```
# plots
```{r}
m_belong_a <- lm(engage ~ belong, d)
plot(m_belong_a) 

m_belong_a_no81 <- lm(engage ~ belong, d_no81)
plot(m_belong_a_no81) 
```
```{r}
coef(m_belong_a)

coef(m_belong_a_no81)
```

# outliers
```{r}
plot(m_belong_a, which = 4)

plot(m_belong_a_no81, which = 4)
```

12. Endorsemnt
# linearity
```{r}
ggplot(d, aes(endorse, engage)) +
geom_point() +
geom_smooth() +
geom_smooth(method = "lm",
color = "magenta")
```
# plots
```{r}
m_endorse_a <- lm(engage ~ endorse, d)
plot(m_endorse_a)

m_endorse_a_no81 <- lm(engage ~ endorse, d_no81)
plot(m_endorse_a_no81)
```

```{r}
coef(m_endorse_a)

coef(m_endorse_a_no81)
```

# outliers
```{r}
plot(m_endorse_a, which = 4)

plot(m_endorse_a_no81, which = 4)
```

13. Value & supp of Ed
# linearity
```{r}
ggplot(d, aes(value_ed, engage)) +
geom_point() +
geom_smooth() +
geom_smooth(method = "lm",
color = "magenta")
```

# plots
```{r}
m_value_ed_a <- lm(engage ~ value_ed, d)
plot(m_value_ed_a)

m_value_ed_a_no81 <- lm(engage ~ value_ed, d_no81)
plot(m_value_ed_a_no81)
```
```{r}
coef(m_value_ed_a)

coef(m_value_ed_a_no81)
```

# outliers
```{r}
plot(m_value_ed_a, which = 4)

plot(m_value_ed_a_no81, which = 4)
```

14. Rel with teachers: comfort, interest, and openness
# linearity
```{r}
ggplot(d, aes(rel_1, engage)) +
geom_point() +
geom_smooth() +
geom_smooth(method = "lm",
color = "magenta")
```
# plots
```{r}
m_rel_1a <- lm(engage ~ rel_1, d)
plot(m_rel_1a)

m_rel_1a_no81 <- lm(engage ~ rel_1, d_no81)
plot(m_rel_1a_no81)
```

```{r}
coef(m_rel_1a)

coef(m_rel_1a_no81)
```

# outliers
```{r}
plot(m_rel_1a, which = 4)

plot(m_rel_1a_no81, which = 4)
```


15. Rel with school: parent self-efficacy to solve probs with school
# linearity
```{r}
ggplot(d, aes(rel_2, engage)) +
geom_point() +
geom_smooth() +
geom_smooth(method = "lm",
color = "magenta")
```

# plots
```{r}
m_rel_2a <- lm(engage ~ rel_2, d)
plot(m_rel_2a)

m_rel_2a_no81 <- lm(engage ~ rel_2, d_no81)
plot(m_rel_2a_no81)
```

```{r}
coef(m_rel_2a)

coef(m_rel_2a_no81)
```

# outliers
```{r}
plot(m_rel_2a, which = 4)

plot(m_rel_2a_no81, which = 4)
```

```{r}
d$q54_p1
```

```{r}
d_no81$q54_p1
```

